---
title: 'Reproducible Research: Peer Assesment1'
author: "Antonio"
date: "8 de abril de 2016"
output: 
    html_document:
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

###1.1 Load Data and preprocessing the data

```{r echo=TRUE}
setwd("C:/Users/Antonio/R/C5")
datos<-read.csv("activity.csv")
summary(datos)
datos$date<-as.Date(datos$date)
```
Find the % number of NA's in data set

```{r echo=TRUE}
mean(is.na(datos$steps))
```

The number is high, we look for a pattern

```{r echo=TRUE}
steps<-datos$steps
nasteps<-is.na(steps)
dias<-datos$date
diasna<-dias[nasteps]
hist(diasna,"day")
```
The NA's are concentrated in 8 sample days we see what day of the week to see if there is repetition, sequence, Maybe data missing every Tuesday? etc

```{r}
unique(sapply(dias[nasteps],weekdays))
```

THe days of the week do not tell us much (not the same day of week, and from Histogram not same week)

###2. What is mean total number of steps taken per day?

2.1 calculate total number of steps per day (NA eliminated)
Histogram of Total Steps Per Day
```{r echo=TRUE}
totalstepsday<-with(datos,tapply(steps,date,sum,na.rm=TRUE))
hist(totalstepsday, main = " Histogram of Steps per Day",xlab = "Total Steps per Day")
abline(v=median(totalstepsday),lwd=3, col="Blue")
abline(v=mean(totalstepsday),lwd=3, col="red")
legend("topright",c("Median","Mean"),col = c("blue","red"),lwd = 3)
```
Boxplot of Total Steps per Day
```{r echo=TRUE}
boxplot(totalstepsday, main = " Steps per Day",ylab = "Total Steps per Day")
```
2.3 Calaculate and report the mean and median, mean of Total Steps per day and summary

```{r echo=TRUE}
summary(totalstepsday)
print(median(totalstepsday))
print(mean(totalstepsday))
```

### 3 What is the average daily activity pattern?
3.1 calculate and Plot the average for time intervals across different dates

```{r echo=TRUE}
avginterv<-with(datos,tapply(steps,interval,mean,na.rm=TRUE))
plot(names(avginterv),avginterv,type = "l",xlab = "Interval", ylab = "Avg. Number Steps per Day", main = "Time Series of Avg. Steps")
```

3.2 Find the 5-minute interval that contains the maximun (average) number of Steps

```{r echo=TRUE}
print(names(which.max(avginterv)))
```
### 4 Imputing missing values
4.1 Calculate and report total number of missing values in dataset

```{r echo=TRUE}
print(sum(is.na(datos$steps)))
```

4.2 Divise a stratgy for missing values in dataset

*4.2.1 I will use the average steps per interval calculated over the dataset days to fill in data

+I create a dataframe with the previously calulated data

+I create an variable holding the positions in which NA data appears

+I create a new dataset which is equal to original data

+Change values for NA in steps variable
```{r echo=TRUE}
avgint<-data.frame(interval=names(avginterv),average=avginterv)
acambiar<-which(nasteps)
datos1<-data.frame(datos)
for (x in acambiar){
    y=datos$interval[x]
    datos1$steps[x]=avgint$average[avgint$interval==y]
}
```
4.4 Make a histogram of total steps taken each day and calculate and report media.

```{r echo=TRUE}
totalstepsday1<-with(datos1,tapply(steps,date,sum,na.rm=TRUE))
hist(totalstepsday1, main = " Histogram of Steps per Day",xlab = "Total Steps per Day")
abline(v=median(totalstepsday1),lwd=6, col="Blue")
abline(v=mean(totalstepsday1),lwd=2, col="red")
legend("topright",c("Median","Mean"),col = c("blue","red"),lwd = 2)
```
And a Boxplot
```{r echo=TRUE}
boxplot(totalstepsday1, main = " Steps per Day",ylab = "Total Steps per Day")
```
Calculate and report the mean and median, mean of Total Steps per day and summary

```{r echo=TRUE}
summary(totalstepsday1)
print(median(totalstepsday1))
print(mean(totalstepsday1))
```
The main impact of using this method of filling data is that the median increases and is now equal to mean. The histogram ### appear to be more "gaussin-normal"

### 5 Are there differences in activity patterns between weekdays and weekends?

5.1 create a new factor variable with two levels - weekeday and weekend

```{r echo=TRUE}
datos1$dayweek<-sapply(datos1$date,weekdays)
days<-unique(datos1$dayweek)
weekday<-days[0:5]
weekend<-days[6:7]
cambio<-function(x){
    if (x %in% weekday){
        x="Weekday"
    } else{
        x="Weekend"
    }
}
datos1$wow<-sapply(datos1$dayweek,cambio)
week<-subset(datos1,datos1$wow=="Weekday")
nweek<-subset(datos1,datos1$wow=="Weekend")
```
5.1 2.	Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis).

```{r echo=TRUE}
avgweek<-with(week,tapply(steps,interval,mean))
avgweekend<-with(nweek,tapply(steps,interval,mean))
davgweek<-data.frame(Interval=as.numeric(names(avgweek)),average=as.numeric(avgweek))
davgweek$day<-"weekday"
davgweekend<-data.frame(Interval=as.numeric(names(avgweekend)),average=as.numeric(avgweekend))
davgweekend$day<-"weekend"
plotsdata<-rbind(davgweek,davgweekend)
library(lattice)
xyplot(average~Interval|day, data = plotsdata, layout=c(1,2), type="l", xlab = "Interval",ylab = "Avg. Number of Steps",xlim = c(0.500,1000,1500,2000,2500))
```







